---
title: "First Look"
output: ioslides_presentation
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<p class="forceBreak"></p>

## The First R Session


```{r, calculator}
6 + 3
5 - 9
4 * 6
8 / 3
5 ^ 2
(1 + 0.05)^8
exp(3)
log(14)
23.76 * log(8)/(23 + atan(9))
```


## assignment

Results of calculations can be stored in objects using the assignment operators:
  
 - an arrow (`<-`) formed by a smaller than character and a hyphen without a space;
 - the equal character (`=`).

These objects can then be used in other calculations. 

There are some restrictions when giving an object a name:
  
 - Object names cannot contain "strange" symbols like `!`, `+`, `-`, `#`.
 - A dot (`.`) and an underscore (`_`) are allowed, also a name starting with a dot.
 - Object names can contain a number but cannot start with a number.
 - R is case sensitive: `X` and `x` are two different objects, as well as `temp` and `temP`.


## assignment

```{r, assignment}
x <- log(14)
y <- 23.76 * log(8)/(23 + atan(9))
z <- x + y
```

## print 

To print the object just enter the name of the object or use `print()` function. 

```{r, print_obj}
x
y
print(z)
```


## Objects listing

Objects created during an R session are hold in your _worspace_ memory. 

`ls()` lists objects

```{r, workspace}
ls()
ls(pattern = 'x')
```


## Removing objects

If a value to an object that already exists is assigned then the contents of the object will be overwritten with the new value (without a warning!). The function `rm()` ought be used to remove one or more objects from your session.

```{r, rm}
friz <- 0
ls()
rm(friz)
ls()
```

## Garbage collection

When objects are no longer used, and this clearly happens when objects are deleted, `R` releases immediately the memory they filled in the system. This is done automatically by the garbage collector `gc()`.

We can call `gc()` also to see how much memory `R` is using for allocating objects

```{r howRworks-029}
gc()
```


## R help

```{r, help1, eval=FALSE}
help(mean)
?sd
```

## R help 

list of functions concerning `mean`

```{r, help.search1, eval=FALSE}
help.search("mean")
```

## R help

list of function containing "mean" in the name 

```{r, help.search2, eval=FALSE}
apropos("mean")
```

