---
title: "2T manipulation"
output: ioslides_presentation
   
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load Tidyverse 

Do not forger to load `tidyverse` and set `options()`

```{r, message = FALSE}
require(tidyverse )
options(tibble.print_max = 5, tibble.print_min = 5)
```


# Mutating Join


## The Basic

Mutating joins allow you to combine variables from multiple tables.  

There are four types of mutating join, which differ in their behavior when a match is not found: 

* `inner_join(x, y)`
* `left_join(x, y)`
* `right_join(x, y)`
* `full_join(x, y)`

## Simple Example 

```{r}
d_1 <- tibble(id = 1:4, x1 = c('a', 'b', 'c', 'd'))
d_2 <- tibble(id = 3:5, x2 = c('c', 'd',  'e'))
```

## inner join

* `inner_join(x, y)` only includes observations that match in both x and y:

```{r}
inner_join(d_1, d_2)
```


## left join 

* `left_join(x, y)` includes all observations in `x`, regardless of whether they match or not. 

This is the most commonly used join because it ensures that you don’t lose observations from your primary table:

```{r}
left_join(d_1, d_2)
```

## right join 

* `right_join(x, y)` includes all observations in `y`. It’s equivalent to `left_join(y, x)`, but the columns will be ordered differently:

```{r}
right_join(d_1, d_2)
```

## full join 

* `full_join()` includes all observations from `x` and `y`:

```{r}
full_join(d_1, d_2)
```


# The Italian Example

## regioni

```{r}
load('./data/regioni.RData')
glimpse (regioni, width = 60)
```


## province

```{r}
load('./data/province.RData')
glimpse (province, width = 60)
```

## comuni

```{r}
load('./data/comuni.RData')
glimpse (comuni, width = 60)
```

## basic join 

```{r}
italia <- comuni %>% 
  left_join(province, by = 'cod_provincia') %>% 
  left_join(regioni, by = 'cod_regione') 
italia
```

## Example

Legal population by regioni sort by descending population 

```{r}
  regioni %>%
  left_join(comuni, by = 'cod_regione') %>%
  group_by(regione) %>%
  summarise(pop_legale = sum(pop_legale, na.rm = T), 
            .groups = 'drop') %>%
  arrange(- pop_legale)
```

Note: na.rm = T



# Filtering Joins

## the basic 

Filtering joins match observations in the same way as mutating joins, but affect observations rather than variables. 

* `semi_join(x, y)` keeps all observations in `x` that have a match in `y`;
* `anti_join(x, y)` drops all observations in `x` that have a match in `y`.



## Simple Example 

```{r}
d_1 <- tibble(id = 1:4, x1 = c('a', 'b', 'c', 'd'))
d_2 <- tibble(id = 3:5, x2 = c('c', 'd',  'e'))
```


All observations in `d_1` and `d_2`

```{r}
d_1 %>% semi_join(d_2)
```

All observations in `d_1` not in `d_2`

```{r}
d_1 %>% anti_join(d_2)
```


## Tennis Example

Data `wimbledon` and `usopen` contain the following variables:

* year
* champion_country
* champion
* runner up country
* runner up


## us open 

```{r}
usopen <- read_rds('./data/usopen.rds')
```


## wimbledon 

```{r}
wimbledon <- read_rds('./data/wimbledon.rds')
```

## acrylic hard courts

who won the US open and never won Wimbledon

```{r}
usopen %>% 
  anti_join(wimbledon, by = "champion") %>%
  select(year, champion_country, champion)
```

## international 

who won both the US open and Wimbledon

```{r}
usopen %>% 
  semi_join(wimbledon, by = "champion") %>%
  select(year, champion_country, champion)
```


## performance

who won both the US open and Wimbledon in the same year


```{r}
usopen %>% 
  semi_join(wimbledon, by = c("champion", "year")) %>%
  select(year, champion_country, champion) 
```

## top performance

who won both the US open and Wimbledon more than once

```{r}
# who won the US open more than once
super_usopen <- usopen %>% 
  group_by(champion) %>%
  summarise (n_us = n(), .groups = 'drop') %>%
  filter(n_us > 1)

# who won Wimbledon more than once
super_wimbledon <- wimbledon %>% 
  group_by(champion) %>%
  summarise (n_wb = n(), .groups = 'drop') %>%
  filter(n_wb > 1)

# semi join
super_usopen %>% 
  semi_join(super_wimbledon, by = "champion") %>%
  select(champion)
```


## best player 

who was the best player in the two slams:

```{r}
super_usopen %>% 
  inner_join(super_wimbledon, by = "champion") %>%
  mutate ( n = n_us+n_wb) %>%
  arrange(desc(n))
```
 


